Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-10 paper.qmd
Update the paper with sumaary for the example, s-curve data set, for s-curve dataset added summary with different parameter setting in UMAP, generate required data sets beforehand and store in necessary folders, generate scripts to generate embeddings

library(Rtsne)
library(umap)
library(phateR)
library(reticulate)
library(patchwork)
library(langevitour)

@@

#### Introduction to UMAP

The UMAP algorithm constructs a fuzzy topological representation of the high-dimensional data and then optimizes the low-dimensional embedding to be as similar to this representation as possible [@Leland2018].

The UMAP algorithm constructs a weighted k-nearest neighbors graph based on the local similarity of data points in the original high-dimensional space. The weighted graph is defined by a conditional probability distribution $P_{ij}$, which represents the probability of observing data point $x_i$ given data point $x_j$ as a neighbor:

$$
P_{ij} = \frac{\exp(-\lVert x_i - x_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert x_i - x_k \rVert^2)}
$$

Similarly, a fuzzy topological representation of the data is constructed in the low-dimensional space with conditional probability distribution $Q_{ij}$:

$$
Q_{ij} = \frac{\exp(-\lVert y_i - y_j \rVert^2)}{\sum_{k \neq i} \exp(-\lVert y_i - y_k \rVert^2)}
$$

where $y_i$ and $y_j$ are the corresponding low-dimensional representations of data points $x_i$ and $x_j$.

UMAP minimizes the cross-entropy between the high-dimensional fuzzy topological representation $P_{ij}$ and the low-dimensional representation $Q_{ij}$ to find an optimal low-dimensional embedding. The loss function for UMAP is defined as:

$$
L = \sum_{i} \sum_{j} P_{ij} \log \frac{P_{ij}}{Q_{ij}}
$$

The algorithm uses stochastic gradient descent to optimize the positions of the data points in the low-dimensional space to minimize this loss function.

In R, the `umap` package is often used to implement UMAP and generate results.

The "n_neighbors" parameter in UMAP determines the number of nearest neighbors to consider for each data point during the construction of the fuzzy topological representation. Higher values of "n_neighbors" capture more global structure, while lower values emphasize local structure. Choosing an appropriate value for "n_neighbors" is crucial, as it affects the balance between local and global structure in the resulting low-dimensional representation.

@@

data <- data |>
  mutate(ID = row_number())

data_split <- initial_split(data)
training_data <- training(data_split) |>
  arrange(ID)
test_data <- testing(data_split) |>
  arrange(ID)
```

@@

tSNE_data <- Fit_tSNE(training_data, opt_perplexity = calculate_effective_perplexity(training_data), with_seed = 20240110)
#(perplexity: ", calculate_effective_perplexity(data), ")
plot_list1 <- plot_tSNE_2D(tSNE_data) + #ggtitle("(a)") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

PHATE_data <- read_csv(paste0(here::here(), "/data/phate_data_s_curve.csv"))

@@

TriMAP_data <- read_csv(paste0(here::here(), "/data/trimap_data_s_curve.csv"))

@@

PaCMAP_data <- read_csv(paste0(here::here(), "/data/pacmap_data_s_curve.csv"))

@@

```{r}
#| echo: false
#| fig-cap: "2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 32), (b) UMAP (n_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2). Is there a best representation of the original data or are they all providing  equivalent information?"
#| label: fig-nldervis
#| out-width: 100%

@@

```{r}
#| echo: false
#| message: false
#| warning: false
hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_8.rds")

@@

```{r}
#| echo: false
#| message: false
#| warning: false
hex_full_count_df_loop3 <- read_rds("data/s_curve/s_curve_hex_15.rds")

@@

```{r}
#| echo: false
#| label: fig-binsize
#| fig-pos: H
#| fig-cap: "Hexbin plots from different number of bins for the same **s_curve_noise_umap** data: (a) b = (4, 8), s = 2.031, (b) b = (9, 20), s = 2.031, and (c) b = (16, 36), s = 2.031. The hexbins are colored based on the density of points, with yellow indicating higher point density and darker colors representing lower point density within each bin. Does  a value of number of bins exist to effectively represent the low-dimensional data?"

@@

## Data set with all possible centroids in the hexagonal grid

full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)

## To map hexID to hexbin centroids in the full grid

vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names

full_grid_with_hexbin_id <- dplyr::bind_rows(vec1)[0, ]
full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
  dplyr::mutate_if(is.character, as.numeric)

for(i in 1:length(sort(unique(full_centroid_df$y)))){

  ## Filter the data set with specific y value
  specific_y_val_df <- full_centroid_df |>
    dplyr::filter(y == sort(unique(full_centroid_df$y))[i])

  ordered_x_df <- specific_y_val_df |>
    dplyr::arrange(x)

  full_grid_with_hexbin_id <- dplyr::bind_rows(full_grid_with_hexbin_id, ordered_x_df)

full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
  dplyr::mutate(hexID = row_number())

full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
  dplyr::rename("c_x" = "x",
         "c_y" = "y")

full_grid_with_hexbin_id <- dplyr::full_join(full_grid_with_hexbin_id, df_bin_centroids, by = c("hexID" = "hexID")) |>
  dplyr::select(-c(x, y)) #|> 
  #dplyr::mutate(counts = tidyr::replace_na(counts, 0))

full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
    dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))

## Generate all coordinates of hexagons
hex_grid <- full_hex_grid(full_centroid_df)

full_grid_with_polygon_id_df <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)

full_grid_with_hexbin_id_rep <- full_grid_with_polygon_id_df |>
  dplyr::slice(rep(1:n(), each = 6)) |>
  dplyr::arrange(polygon_id)

hex_full_count_df <- dplyr::bind_cols(hex_grid, full_grid_with_hexbin_id_rep)

@@

@@

```{r}
#| echo: false
predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_predict.rds")
```

predict_umap_plot <- plot_UMAP_2D(UMAP_data) +
  geom_point(data = predict_UMAP_df, aes(x = UMAP1, y = UMAP2), color = "red") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)
```

<!--predict hexid for the test data -->

pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_x, shape_val = shape_value)
pred_df_training <- pred_df_training_object$pred_data
centroid_df_training <- pred_df_training_object$df_bin_centroids
avg_df_training <- pred_df_training_object$df_bin

eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_x)

pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_x, shape_val = shape_value)
pred_df_test <- pred_df_test_object$pred_data
centroid_df_test <- pred_df_test_object$df_bin_centroids
avg_df_test <- pred_df_test_object$df_bin

eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_x)
```

## Data set with all possible centroids in the hexagonal grid
full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)

## Generate all coordinates of hexagons
hex_grid <- full_hex_grid(full_centroid_df)

#hex_full_count_df <- generate_full_grid_info(df_bin_centroids)

predict_hex_ID_plot <- ggplot(data = hex_grid, aes(x = x, y = y)) + geom_polygon(fill = "white", color = "black", aes(group = id)) +
  geom_point(data = df_bin_centroids, aes(x = x, y = y), color = "#33a02c") +
  geom_point(data = predict_UMAP_df, aes(x = UMAP1, y = UMAP2), color = "red") +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
    theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3) 
```

<!--To draw predict values in hexbin grid-->

```{r}
#| echo: false
#| fig-cap: The .
#| label: fig-predict

predict_umap_plot  + predict_hex_ID_plot +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

```{r}
#| echo: false
#num_bins_vec <- append(sample(1:sqrt(750), 10), num_bins_x)

#num_bins_vec <- sample(2:1000, 40)
num_bins_vec <- 1:10

@@

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value)
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i])

  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value)
  pred_df_test <- pred_df_test_object$pred_data
  centroid_df_test <- pred_df_test_object$df_bin_centroids
  avg_df_test <- pred_df_test_object$df_bin

  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i])

@@

}
```

@@

MSE_df <- dplyr::bind_rows(eval_data_training, eval_data_test)
```

```{r}
#| echo: false
aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                   y = total_error,
                   color = data_type
)) +  
  geom_point() +
  geom_line() +
  #geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-15000, label=paste0("effective number of bins = ",as.character(NROW(full_grid_with_hexbin_id))), angle=90, size = 2.5) + 
  theme(legend.position = "none", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
          axis.title = element_text(size = 7),
        legend.text = element_text(size=7)) +
  scale_color_manual(values = c("#d95f02")) + 
  ylab("AIC") +
  xlab("Total number of bins")
```

```{r}
#| echo: false
mse_plot <- ggplot(MSE_df, aes(x = number_of_bins,
                   y = total_mse,
                   color = data_type
)) +  
  geom_point() +
  geom_line() +
  #geom_smooth(method = "loess", se = FALSE) +
  geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.05, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90, size = 2.5) +
  scale_color_manual(values = c("#1b9e77", "#d95f02")) + 
  theme(legend.position = "bottom", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
          axis.title = element_text(size = 7),
        legend.text = element_text(size=7)) + 
  ylab("MSE") +
  xlab("Total number of bins")
```

aic_plot + mse_plot +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect') &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

>>>>

library(Rtsne)
library(umap)
library(phateR)
library(patchwork)
library(langevitour)
library(colorspace)

@@

#### Non-linear dimension reduction techniques

Non-linear dimension reduction techniques play a crucial role in the analysis and visualization of high-dimensional data, where the complexities of relationships among variables may not be adequately captured by linear methods. Among these techniques, tSNE stands out for its emphasis on preserving pairwise distances. By minimizing the divergence between probability distributions in both the high and low-dimensional spaces, t-SNE effectively reveals intricate structures and patterns within the data. Its application is widespread in tasks requiring the visualization of clusters and local relationships, though it does require careful consideration of the perplexity parameter for optimal results.

UMAP is another powerful non-linear technique that strikes a balance between preserving local and global structures. Constructing a fuzzy topological representation using a weighted k-nearest neighbors graph, UMAP optimizes the low-dimensional embedding to resemble this representation. Known for its efficiency and scalability, UMAP is versatile across various scales of relationships in the data, although parameter sensitivity, particularly concerning the choice of neighbors, must be taken into account.

For trajectory data, PHATE provides specialized capabilities. It models the affinity between data points, simulating a heat diffusion process to capture developmental processes, particularly in single-cell genomics. While PHATE excels in revealing trajectory structures and offering insights into cellular development, it necessitates careful parameter tuning due to its specialized nature.

TriMAP adopts a unique approach by approximating the data manifold through the construction of a triangulated graph representation. This technique efficiently captures both global and local structures by representing the data as a network of triangles. TriMAP's strength lies in its ability to efficiently capture complex structures, albeit with sensitivity to parameter choices, including the number of neighbors.

In contrast, PaCMAP introduces supervised learning to create a low-dimensional representation while preserving pair-wise relationships. Constructing a graph based on pair-wise distances, PaCMAP optimizes an embedding using a customizable loss function. Particularly notable is PaCMAP's flexibility in incorporating class labels or additional information to guide the embedding process, offering users a means to customize its behavior and performance.

@@

# data <- data |>
#   mutate(ID = row_number())
# 
# data_split <- initial_split(data)
# training_data <- training(data_split) |>
#   arrange(ID)
# test_data <- testing(data_split) |>
#   arrange(ID)
training_data <- read_rds("data/s_curve/s_curve_training.rds")
test_data <- read_rds("data/s_curve/s_curve_test.rds")
```

@@

tSNE_data <- read_rds("data/s_curve/s_curve_tsne_27.rds")
#(perplexity: ", calculate_effective_perplexity(data), ")
plot_list1 <- plot_tSNE_2D(tSNE_data) + #ggtitle("(a)") + 
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

PHATE_data <- read_rds(file = "data/s_curve/s_curve_phate.rds")

@@

TriMAP_data <- read_rds(file = "data/s_curve/s_curve_trimap.rds")

@@

PaCMAP_data <- read_rds(file = "data/s_curve/s_curve_pacmap.rds")

@@

```{r}
#| echo: false
#| fig-cap: "2D layouts from different NLDR techniques applied the same data: (a) tSNE (perplexity = 27), (b) UMAP (n_neighbors = 50), (c) PHATE (knn = 5), (d) TriMAP (n_inliers = 5, n_outliers = 4, n_random = 3), and (e) PaCMAP (n_neighbors = 10, init = random, MN_ratio = 0.9, FP_ratio = 2). Is there a best representation of the original data or are they all providing  equivalent information?"
#| label: fig-nldervis
#| out-width: 100%

@@

```{r}
#| echo: false
#| message: false
#| warning: false
hex_full_count_df_loop2 <- read_rds("data/s_curve/s_curve_hex_11.rds")

@@

```{r}
#| echo: false
#| message: false
#| warning: false
hex_full_count_df_loop3 <- read_rds("data/s_curve/s_curve_hex_20.rds")

@@

```{r}
#| echo: false
#| label: fig-binsize
#| fig-pos: H
#| fig-cap: "Hexbin plots from different number of bins for the same **s_curve_noise_umap** data: (a) b = (4, 8), s = 1.643542, (b) b = (12, 22), s = 1.643542, and (c) b = (21, 40), s = 1.643542. The hexbins are colored based on the density of points, with yellow indicating higher point density and darker colors representing lower point density within each bin. Does  a value of number of bins exist to effectively represent the low-dimensional data?"

@@

# ##########
# 
# ## Data set with all possible centroids in the hexagonal grid
# 
# full_centroid_df <- generate_full_grid_centroids(df_bin_centroids)
# 
# ## To map hexID to hexbin centroids in the full grid
# 
# vec1 <- stats::setNames(rep("", 2), c("x", "y"))  ## Define column names
# 
# full_grid_with_hexbin_id <- dplyr::bind_rows(vec1)[0, ]
# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
#   dplyr::mutate_if(is.character, as.numeric)
# 
# for(i in 1:length(sort(unique(full_centroid_df$y)))){
#   
#   ## Filter the data set with specific y value
#   specific_y_val_df <- full_centroid_df |>
#     dplyr::filter(y == sort(unique(full_centroid_df$y))[i])
#   
#   ordered_x_df <- specific_y_val_df |>
#     dplyr::arrange(x) 
#   
#   full_grid_with_hexbin_id <- dplyr::bind_rows(full_grid_with_hexbin_id, ordered_x_df)
#   
# }
# 
# 
# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
#   dplyr::mutate(hexID = row_number())
# 
# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
#   dplyr::rename("c_x" = "x",
#          "c_y" = "y") 
# 
# full_grid_with_hexbin_id <- dplyr::full_join(full_grid_with_hexbin_id, df_bin_centroids, by = c("hexID" = "hexID")) |>
#   dplyr::select(-c(x, y)) #|> 
#   #dplyr::mutate(counts = tidyr::replace_na(counts, 0))
# 
# full_grid_with_hexbin_id <- full_grid_with_hexbin_id |>
#     dplyr::mutate(std_counts = counts/max(counts, na.rm = TRUE))
# 
# ## Generate all coordinates of hexagons
# hex_grid <- full_hex_grid(full_centroid_df)
# 
# full_grid_with_polygon_id_df <- map_polygon_id(full_grid_with_hexbin_id, hex_grid)
# 
# full_grid_with_hexbin_id_rep <- full_grid_with_polygon_id_df |>
#   dplyr::slice(rep(1:n(), each = 6)) |>
#   dplyr::arrange(polygon_id)

hex_full_count_df <- read_rds("data/s_curve/s_curve_hex_11.rds")

@@

In our model tuning process, we strategically adjust three key parameters to optimize the performance and accuracy of our approach. They are (i) the total number of bins ($b$), (ii) a benchmark value to remove low-density hexagons, and (iii) a benchmark value to remove long edges.

### Total number of bins

The total number of bins is a crucial parameter that defines the granularity of our hexagonal grid. This parameter particularly influential in capturing the intricacies of the data structure. It is determined by multiplying the number of bins along the x-axis (\(b_1\)) with the number of bins along the y-axis (\(b_2\)), as per the formula:

$$
b = b_1 \times b_2
$${#eq-equationp1}

Here, \(b\) represents the total number of bins. Adjusting the parameter \(b_1\) provides control over the total bin count (\(b\)), allowing us to fine-tune the grid's configuration along the x-axis based on specific requirements.

### Benchmark value to remove low-density hexagons

Addressing low-density hexagons is a systematic process to handle sparsely represented data in certain regions. For each hex bin, we identify the six nearest hex bins using an equal 2D distance metric. Then, we calculate the mean density, as outlined in the equations:

$$
\text{standard count} = \frac{\text{count}}{\text{max count}} 
$${#eq-equationp2}

$$
\text{mean density} = \frac{\text{standard count}}{6} 
$${#eq-equationp3}

The standard count is derived from the number of observations in the hex bins. By examining the distribution of mean densities and designating the first quartile as the benchmark value, hex bins with mean densities below this benchmark are removed. This process ensures the elimination of regions with insufficient data density, focusing on areas with more significant data representation and preserving the overall structure in the low-dimensional space.

### Benchmark value to remove long edges

The removal of long edges is a critical step to create a smoother representation  by iteratively eliminating hexagons with excessive distances between centroids. This process eliminates outliers and noise while preserving essential local relationships within the data. To achieve this, distances between vertices are sorted, and unique distance values are extracted. By setting a threshold based on the largest difference between consecutive distance values, long edges are identified and removed. This refinement step contributes to enhancing the quality of the triangular mesh, ensuring a more accurate representation of the data structure.

@@

data <- read_csv("data/s_curve.csv")
data <- data |>
  mutate(ID = row_number())

## tSNE

shape_value_curve <- calculate_effective_shape_value(.data = tSNE_data,
                                                   x = tSNE1, y = tSNE2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = tSNE_data, nldr_df_test = tSNE_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "tSNE1", y = "tSNE2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)

## Add new column with data types

MSE_df_1 <- eval_data_training |>
  dplyr::mutate(method = "tSNE")
```

## UMAP
## Prediction

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
                                                   x = UMAP1, y = UMAP2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)

## Add new column with data types

MSE_df_2 <- eval_data_training |>
  dplyr::mutate(method = "UMAP")
```

```{r}
#| warning: false
#| echo: false
## PAHTE
## Prediction

shape_value_curve <- calculate_effective_shape_value(.data = PHATE_data,
                                                   x = PHATE1, y = PHATE2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = PHATE_data, nldr_df_test = PHATE_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "PHATE1", y = "PHATE2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)

## Add new column with data types

MSE_df_3 <- eval_data_training |>
  dplyr::mutate(method = "PHATE")

## TriMAP

## Prediction

shape_value_curve <- calculate_effective_shape_value(.data = TriMAP_data,
                                                   x = TriMAP1, y = TriMAP2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = TriMAP_data, nldr_df_test = TriMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "TriMAP1", y = "TriMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)

## Add new column with data types

MSE_df_4 <- eval_data_training |>
  dplyr::mutate(method = "TriMAP")

## PaCMAP

## Prediction

shape_value_curve <- calculate_effective_shape_value(.data = PaCMAP_data,
                                                   x = PaCMAP1, y = PaCMAP2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = PaCMAP_data, nldr_df_test = PaCMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "PaCMAP1", y = "PaCMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  eval_data_training <- dplyr::bind_rows(eval_data_training, eval_df_training)

## Add new column with data types

MSE_df_5 <- eval_data_training |>
  dplyr::mutate(method = "PaCMAP")

MSE_df <- dplyr::bind_rows(MSE_df_1, MSE_df_2, MSE_df_3, MSE_df_4, MSE_df_5)

## To draw with AIC
aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                                                                                 y = total_error,
                                                                                 color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_color_discrete_qualitative() +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_plot <- ggplot(MSE_df, aes(x = number_of_bins,
                                       y = total_mse,
                                       color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

aic_plot + mse_plot +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

## UMAP with n_neighbor: 7
## Prediction

UMAP_data <- read_rds("data/s_curve/s_curve_umap_7.rds")
predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_7_predict.rds")

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
                                                     x = UMAP1, y = UMAP2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

@@

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_test <- pred_df_test_object$pred_data
  centroid_df_test <- pred_df_test_object$df_bin_centroids
  avg_df_test <- pred_df_test_object$df_bin

  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")

@@

## Add new column with data types

@@

MSE_df_umap_1 <- bind_rows(eval_data_training, eval_data_test) |>
  dplyr::mutate(param = "n_neighbors: 7")
```

## UMAP with n_neighbor: 15
## Prediction

UMAP_data <- read_rds("data/s_curve/s_curve_umap_15.rds")
predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_15_predict.rds")

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
                                                     x = UMAP1, y = UMAP2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_test <- pred_df_test_object$pred_data
  centroid_df_test <- pred_df_test_object$df_bin_centroids
  avg_df_test <- pred_df_test_object$df_bin

  eval_df_test <- generate_eval_df(data = data, prediction_df = pred_df_test, df_bin_centroids = centroid_df_test, df_bin = avg_df_test, num_bins = num_bins_vec[i], col_start = "x")

## Add new column with data types

MSE_df_umap_2 <- bind_rows(eval_data_training, eval_data_test) |>
  dplyr::mutate(param = "n_neighbors: 15")
```

## UMAP with n_neighbor: 32
## Prediction

UMAP_data <- read_rds("data/s_curve/s_curve_umap_32.rds")
predict_UMAP_df <- read_rds("data/s_curve/s_curve_umap_32_predict.rds")

shape_value_curve <- calculate_effective_shape_value(.data = UMAP_data,
                                                     x = UMAP1, y = UMAP2)

num_bins_vec <- 1:10 ## Number of bins along the x-axis

  pred_df_training_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = UMAP_data, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_training <- pred_df_training_object$pred_data
  centroid_df_training <- pred_df_training_object$df_bin_centroids
  avg_df_training <- pred_df_training_object$df_bin

  eval_df_training <- generate_eval_df(data = data, prediction_df = pred_df_training, df_bin_centroids = centroid_df_training, df_bin = avg_df_training, num_bins = num_bins_vec[i], col_start = "x")

  pred_df_test_object <- predict_hex_id(training_data = training_data, nldr_df = UMAP_data, nldr_df_test = predict_UMAP_df, num_bins = num_bins_vec[i], shape_val = shape_value_curve, x = "UMAP1", y = "UMAP2", col_start = "x")
  pred_df_test <- pred_df_test_object$pred_data
  centroid_df_test <- pred_df_test_object$df_bin_centroids
  avg_df_test <- pred_df_test_object$df_bin
